---
title: "Tragically, not Texas"
author: "Leah N. Crowley"
date: "2023-11-14"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Call relevant libraries to workspace for this assignment: 
  library(biscale)
  library(car)
  library(cleangeo)
  library(cowplot)
  library(geosphere)
  library(maps)
  library(mapproj)
  library(maptools)
  library(spatialreg)
  library(spdep)
  library(tidyverse)
  library(visreg)

```


```{r data, include=FALSE}

data <- read.csv("childpov18_southfull.csv")
LA   <- data %>% filter(State == "LA")

names(LA)[names(LA)=="X2016.child.poverty"] <- "child.pov.2016"

```

# Ordinary Least Squares 

Test the relationship of unemployment in Louisiana and the other variables in the dataset: 

```{r OLS, echo=FALSE, message=FALSE, warning=FALSE}

equation <- unemployment_13 ~ rural + urban + lnmanufacturing + lnag + 
  lnretail + lnhealthss + lnconstruction + lnlesshs + 
  lnchildpov + lnsinglemom + lnblack + lnhispanic + 
  lnuninsured + lnincome_ratio + lnteenbirth + lnunmarried

options(scipen = 5)

ols <- lm(equation, data=LA)
summary(ols)

```

```{r OLS Visual, echo=FALSE, message=FALSE, warning=FALSE}

par(mfrow=c(4,4), mar=c(2, 2, 2, 2))
visreg(ols)

```

# Contiguity (Spatial Regression Analysis)

## Create a list of contiguity neighbors: 
```{r Contiguity Neighbors, echo=FALSE, message=FALSE, warning=FALSE}

fips <- county.fips
fips.codes <- separate(data = fips, col = polyname, into = c("state", "county"), sep = ",")
LA_fips <- subset(fips.codes, state=="louisiana", select=fips)

louisiana <- maps::map(database = "county", regions = "louisiana", fill=T, plot=F)
LA_sp = map2SpatialPolygons(louisiana, LA_fips$fips, CRS("+proj=longlat"))

cleaned <- clgeo_Clean(LA_sp)
neighb.data <- poly2nb(cleaned, queen=T)
cont.neighb <- nb2listw(neighb.data,style="W", zero.policy = TRUE)

```

## Moran’s Correlation and LaGrange Multiplier Tests
```{r Moran and LaGrange, echo=FALSE, message=FALSE, warning=FALSE}

lm.morantest(ols, cont.neighb)

lm.LMtests(ols, cont.neighb, test="all")

```

## Spatially lagged X model
```{r Lagged, echo=FALSE, message=FALSE, warning=FALSE}

SLX.model <- spatialreg::lmSLX(equation, data=LA, cont.neighb)
summary(SLX.model)

summary(spatialreg::impacts(SLX.model, cont.neighb), zstats = TRUE)[["pzmat"]]

```

## Spatial lag model
```{r Lag, echo=FALSE, message=FALSE, warning=FALSE}

sp.lag.model <- spatialreg::lagsarlm(equation, data=LA, cont.neighb)
summary(sp.lag.model, Nagelkerke = TRUE)

summary(spatialreg::impacts(sp.lag.model, listw = cont.neighb, R=100), zstats = TRUE)[["pzmat"]]

```
## Spatial error model
```{r Spatial error, echo=FALSE, message=FALSE, warning=FALSE}

sp.err.model <- spatialreg::errorsarlm(equation, data=LA, cont.neighb)
summary(sp.err.model, Nagelkerke = TRUE)

spatialreg::Hausman.test(sp.err.model)

```

## Spatial Durbin model
```{r Durbin, echo=FALSE, message=FALSE, warning=FALSE}

sd.err <- spatialreg::errorsarlm(equation, LA, cont.neighb, etype = "emixed")
sdm <- spatialreg::lagsarlm(equation, LA, cont.neighb, type = "mixed")

summary(sd.err, Nagelkerke = TRUE)

```

# K-nearest neighbors 
## Create list of k-neighbors 
```{r k-neighbors, echo=FALSE, message=FALSE, warning=FALSE}

all.xy <-centroid(LA_sp)
colnames(all.xy) <- c("x","y")

# Create neighbors
  all.dist.k1 <- knn2nb(knearneigh(all.xy, k=1, longlat = TRUE))
  all.dist.k3 <- knn2nb(knearneigh(all.xy, k=3, longlat = TRUE))
  all.dist.k5 <- knn2nb(knearneigh(all.xy, k=5, longlat = TRUE))

# Determine max k distance value to neighbor
  all.max.k1 <- max(unlist(nbdists(all.dist.k1, all.xy, longlat=TRUE)))
  all.max.k3 <- max(unlist(nbdists(all.dist.k3, all.xy, longlat=TRUE)))
  all.max.k5 <- max(unlist(nbdists(all.dist.k5, all.xy, longlat=TRUE)))

# Calculate neighbors based on distance
  all.sp.dist.k1 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k1, longlat = TRUE)
  all.sp.dist.k3 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k3, longlat = TRUE)
  all.sp.dist.k5 <- dnearneigh(all.xy, d1=0, d2=1 * all.max.k5, longlat = TRUE)

# Create neighbor list
  all.dist.neighb.k1 <- nb2listw(all.sp.dist.k1,style="W", zero.policy = TRUE)
  all.dist.neighb.k3 <- nb2listw(all.sp.dist.k3,style="W", zero.policy = TRUE)
  all.dist.neighb.k5 <- nb2listw(all.sp.dist.k5,style="W", zero.policy = TRUE)
  
```

## Calculate distance lag model for each k-distance: 
```{r k lag model, echo=FALSE, message=FALSE, error=FALSE}

all.dist.lag.k1 <- spatialreg::lagsarlm(equation, data = LA, listw = all.dist.neighb.k1)
all.dist.lag.k3 <- spatialreg::lagsarlm(equation, data = LA, listw = all.dist.neighb.k3)
all.dist.lag.k5 <- spatialreg::lagsarlm(equation, data = LA, listw = all.dist.neighb.k5)

summary(all.dist.lag.k1, Nagelkerke = TRUE)

```

## Create distance error model for each k-distance: 
```{r k error model, echo=FALSE, message=FALSE, error=FALSE}

all.dist.err.k1 <- spatialreg::errorsarlm(equation, data = LA, listw = all.dist.neighb.k1)
all.dist.err.k3 <- spatialreg::errorsarlm(equation, data = LA, listw = all.dist.neighb.k3)
all.dist.err.k5 <- spatialreg::errorsarlm(equation, data = LA, listw = all.dist.neighb.k5)

summary(all.dist.err.k1, Nagelkerke = TRUE)

```

# Map time! 
## Combine columns from unemployment and error model data sets:
```{r mapping, echo=FALSE, message=FALSE, error=FALSE}

dist.err.data <- summary(all.dist.err.k1, correlation=TRUE, Nagelkerke = TRUE)

dist.err.output <- cbind.data.frame(LA$FIPS,
                                    dist.err.data$fitted.values, 
                                    dist.err.data$residual, 
                                    LA$child.pov.2016, 
                                    LA$lnsinglemom, 
                                    LA$lnuninsured, 
                                    LA$lnlesshs, 
                                    LA$lnunemployment,
                                    LA$lnincome_ratio,
                                    stringsAsFactors = FALSE)

# Renaming columns
  colnames(dist.err.output) <- c("fips","fitted","resid","childpov",
                          "single_mom","uninsured","less_hs","unemployed","income_ratio")

```


```{r fortify, error=FALSE, message=FALSE, include=FALSE}

LA_fortify <- fortify(LA_sp)

LA_poly <- merge(x = LA_fortify, y = dist.err.output, 
                 by.x = "id", by.y = "fips", all = TRUE)

bivariate_data <- bi_class(LA_poly, x = uninsured, y = unemployed, 
                           dim = 3, style = "quantile")

legend <- bi_legend(pal = "PinkGrn",
                    dim = 4,
                    xlab = "Uninsured Individuals",
                    ylab = "Unemployed Individuals",
                    size = 6)

```

```{r basemap, include=FALSE}

world <- map_data("world")
states <- map_data("state")
southern_states <- subset(states, region %in% 
                            c("texas", "arkansas", "louisiana", "mississippi", 
                              "alabama", "georgia", "florida", "north carolina",
                              "south carolina", "tennessee", "oklahoma", 
                              "kentucky", "west virginia", "virginia", 
                              "maryland", "delaware", "district of columbia"))

```

```{r map, echo=FALSE, message=FALSE, warning=FALSE}

mom_pov_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long,y=lat, group=group), fill = "gray95", color = "black") +
  geom_polygon(data = states, aes(x=long,y=lat, group=group), fill = "lightgray", color = "black") +
  geom_polygon(data = southern_states, aes(x=long,y=lat, group=group), fill = NA, size = 0.01, color = "black") +  
  geom_polygon(data = bivariate_data, aes(x=long, y=lat, group=group, fill = bi_class), color = "black", show.legend = FALSE) + 
  bi_scale_fill(pal = "PinkGrn", dim = 3) +
  coord_map("conic", lat0 = 30, xlim=c(-95.5,-88.4), ylim=c(26,34)) +
  theme_void() + theme(legend.title.align=0.5) +
  theme(panel.background = element_rect(fill = 'lightblue'),
        panel.grid.major = element_line(colour = NA)) +
  labs(x = "Longitude", y = "Latitude", fill = "Child Poverty", 
       title = "Unemployment and Uninsured Individuals in LA") +
  theme(plot.title = element_text(face = "bold", hjust = 0.25))

final_map <- ggdraw() +
  draw_plot(mom_pov_map, 0, 0, 1, 1) +
  draw_plot(legend, 0.375, 0.05, 0.25, 0.25)

final_map

```